---
title: "R Notebook"
output: html_notebook
---
```{r}
#rm(list = ls())
setwd("C:/Users/LiaDenisse/Documents/OR and DS/TSP")
library(Rcpp)
#Rcpp::sourceCpp("OR and DS/TSP/functionsC.cpp")
```

TSP problem

Define useful functions
```{r}

distance_att <- function(df, i, j){
  f1 <- as.numeric(df[i,][1]) - as.numeric(df[j,][1])
  f2 <- as.numeric(df[i,][2]) - as.numeric(df[j,][2])
  xdist <- ((f1*f1) + (f2*f2))/10
  xdist <- sqrt(xdist)
  tdist <- round(xdist)
  if(tdist < xdist) {dist <- tdist + 1}
  else {dist <- tdist}
  return(dist)
}


distance_inf <- function(tsp, point1, point2){
  if(tsp[point1,point2] == 0) { return(1000)} else{
    if(tsp[point1,point2] == 0.0001) {return(0)} else{
       return(tsp[point1,point2])}}
}

distance <- function(tsp, point1, point2){
       return(tsp[point1,point2])
}

tour_to_df <- function(dist_matrix, tour){
  n <- length(tour)
  df <- data.frame(matrix(ncol = 3, nrow = (n-1)))
  colnames(df) <- c('from', 'to', 'distance')
  for(i in 1:(n-1)){
    df$from[i] <- tour[i]
    df$to[i] <- tour[i+1]
    df$distance[i] <- dist_matrix[tour[i], tour[i+1]]
  }
  return(df)
}

path_length <- function(tsp, path){
  if (length(path) == 1) { return (0)}
  else {
    start_node <- path[1]
    next_node <- path[-1][1]
    path <- path[-1]
    return (distance(tsp,start_node,next_node)+path_length(tsp,path))
  }
}

#tour_length( distance matrix, tour) C++ function

#nearest_neighbor( distance matrix, tour) C++ function

furthest_neighbor <- function(tsp,  current_city, untraveled_cities) {
  fn <- 0
  max_dist <- 0
  for (city in untraveled_cities) {
    if (distance(tsp,current_city,city) > max_dist) {
      fn <- city
      max_dist <- distance(tsp,current_city,city)
    } 
  }
  return (fn)
}

#return position (i,j) of minimum element in a matrix
min_matrix <- function(M){
  return (arrayInd(which.min(M),dim(M)))
}

# total time: time_matrix a matrix, collection and tour vectors
tour_time <- function(time_matrix, collection, tour){
  traveling_time <- path_length(time_matrix, tour)
  collection_time <- sum(collection[tour])
  return(traveling_time + collection_time)
}

#reorder a tour so that it starts from node "start"
reorder_tour <- function(tour,start){
  if(tour[1] == start){ ordered_tour <- tour} 
  else {
 ordered_tour <- c( tour[which(tour==start):length(tour)], tour[2:(which(tour==start))])}
  
 return(ordered_tour)
}

# try to find a tour using the nodes in the given cluster, by trying different construction heuristcs and improvement heuristics. Output: best tour found (the more feasible tour, meaning, the one with the least number of forbidden edges) + TRUE or FALSE stating whether the resulting tour is feasible.



is_initial_tour_feasible <- function(solution, cluster, first, M, dumpsite){
  if(missing(first)){first <- FALSE}

  if(first) {
    st <- c(1,solution[[cluster]],dumpsite)
    st<- st[order(st)]
    M1 <- M[st,st]
  } else {
    st <- c(dumpsite,solution[[cluster]])
    st<- st[order(st)]
    M1 <- M[st,st]
  }
  
  tour_options <- list()
  
  if(first){
  nn1<- nearest_neighbor_tour_start(M1,1) 
  tour_options[[1]] <- st[nn1]
  } else {
  
  nn1<- nearest_neighbor_tour_start(M1,dim(M1)[1]) 
  tour_options[[1]] <- st[nn1]  
  }
  
  gr1<- greedy_algorithm_tour(M1) 
  tour_options[[2]] <- st[gr1]
  
  ni1<- greedy_algorithm_tour(M1) 
  tour_options[[3]] <- st[ni1]
  
   min <- Inf
  for(i in 1:3){
    if( tour_length(M,tour_options[[i]])< min){
      min <- tour_length(M,tour_options[[i]])
      choosen_tour <- tour_options[[i]]
    }
  }
  
  i2opt <- improvement_2opt(M,choosen_tour, asymmetric = TRUE)
  i3opt <- ThreeOpt_R_greedy(i2opt,M,max_improvements = 50)
  
  if(first){
    tour <- reorder_tour(i3opt,1)} else {
    tour <- reorder_tour(i3opt,dumpsite)
  }
  
  res <- list()
  res[[1]] <- tour
  if(tour_length(M,tour) < 1000) {
    res[[2]] <- TRUE
  } else {
    res[[2]] <- FALSE
  }
  
  return(res)

}

# if no feasible tour is found, try this. This will add nodes to the initial set of nodes until a feasible tour can be constructed by using nn algorithm. 




make_tour_feasible <- function(dist_matrix, set_of_nodes, first){
  continuar <- TRUE
  nadd <- 0
  M_size <- dim(dist_matrix)[1]
  
  while(continuar){
  
  add <- nearest_insertion(dist_matrix,set_of_nodes)
  nadd <- nadd + 1
  set_of_nodes <- c(set_of_nodes,add)
  set_of_nodes<- set_of_nodes[order(set_of_nodes)]
  M1 <- dist_matrix[set_of_nodes,set_of_nodes]
  
  nn1<- nearest_neighbor_tour_start(M1,1) 
  nn <- set_of_nodes[nn1]
  tour_length(dist_matrix,nn) 
  
  if ((tour_length(dist_matrix,nn) < 1000) || (length(unique(nn)) == M_size )) { continuar <- FALSE}
  }
    return(nn)
}


make_tour_feasible2 <- function(dist_matrix, set_of_nodes, dumpsite, first, centroid){
  continuar <- TRUE
  nadd <- 0
  M_size <- dim(dist_matrix)[1]
  all_nodes <- seq(1:M_size)
  other_nodes <- all_nodes[!all_nodes %in% c(set_of_nodes,1,dumpsite)]
  
  if(first) { other_nodes <- c(other_nodes,1)} else {
    other_nodes <- c(other_nodes,dumpsite)
  }
  
  while(continuar){
  
  add <- nearest_neighbor(dist_matrix,centroid,other_nodes)
  nadd <- nadd + 1
  set_of_nodes <- c(set_of_nodes,add)
  set_of_nodes<- set_of_nodes[order(set_of_nodes)]
  M1 <- dist_matrix[set_of_nodes,set_of_nodes]
  
  if(first) {  nn1<- nearest_neighbor_tour_start(M1,1) } else{
    nn1<- nearest_neighbor_tour_start(M1,dumpsite)
  }
  nn <- set_of_nodes[nn1]
  tour_length(dist_matrix,nn) 
  
  if ((tour_length(dist_matrix,nn) < 1000) || (length(unique(nn)) == M_size )) { continuar <- FALSE}
  }
    return(nn)
}



```


Nearest neighbor tour: constructs a tour through all nodes in a TSP by following the nearest neighbor heuristic
```{r}
nearest_neighbor_tour_start<- function(tsp, start){
  
  nn_path <- start
  current_node <- start
  all_nodes <- seq(1,dim(tsp)[1],1) #list all nodes
  nodes_to_visit <- all_nodes[!(all_nodes %in% start)] #all nodes except start node
  
  while (length(nodes_to_visit)>0) {
    current_node <- nearest_neighbor(tsp,current_node,nodes_to_visit)
    nn_path <- c(nn_path,current_node)
    nodes_to_visit <- nodes_to_visit[!(nodes_to_visit %in% current_node)] 
  }
  return (c(nn_path,start))
}

repetitive_nearest_neighbor_tour <- function(tsp){
  
  all_nodes <- seq(1,dim(tsp)[1],1) #list all nodes
  min_length <- Inf
  tour <- 0
  for (node in all_nodes){
    #current_length <- path_length(tsp,nearest_neighbor_tour_start(tsp,node))
    current_length <- tour_length(tsp,nearest_neighbor_tour_start(tsp,node))
    if(current_length < min_length) {
      tour <- nearest_neighbor_tour_start(tsp,node)
      min_length <- current_length }
  }
  return (tour)
}

```
Furthest neighbor tour: constructs a tour through all nodes in a TSP by following the furthest neighbor heuristic
```{r}
furthest_neighbor_tour_start<- function(tsp, start){
  
  nn_path <- start
  current_node <- start
  all_nodes <- seq(1,dim(tsp)[1],1) #list all nodes
  nodes_to_visit <- all_nodes[!(all_nodes %in% start)] #all nodes except start node
  
  while (length(nodes_to_visit)>0) {
    current_node <- furthest_neighbor(tsp,current_node,nodes_to_visit)
    nn_path <- c(nn_path,current_node)
    nodes_to_visit <- nodes_to_visit[!(nodes_to_visit %in% current_node)] 
  }
  return (c(nn_path,start))
}
  
repetitive_furthest_neighbor_tour <- function(tsp){
  
  all_nodes <- seq(1,dim(tsp)[1],1) #list all nodes
  min_length <- Inf
  tour <- 0
  for (node in all_nodes){
    #current_length <- path_length(tsp,furthest_neighbor_tour_start(tsp,node))
     current_length <- tour_length(tsp,furthest_neighbor_tour_start(tsp,node))
    if(current_length < min_length) {
      tour <- furthest_neighbor_tour_start(tsp,node)
      min_length <- current_length }
  }
  return (tour)
}

```

Insertion algorithms. Start with a tour consisting of two nodes and then choose in each step a node k not yet on the tour. This node is inserted into the existing tour between two consecutive nodes i, j such that the insertion cost is minimized. The insertion algorithms differ in the way the node to be inserted is chosen:
* Nearest insertion.

```{r}
nearest_insertion_tour <- function(tsp,current_tour, max_length){
  
  if(missing(max_length)){ max_length <- Inf }
  
  if(missing(current_tour)) {
    n <- dim(tsp)[1]
    C<- tsp  # distances matrix
    for (i in 1:n){
      C[i,i] <- Inf
    }
     current_tour <- min_matrix(C)
  }
  
  
  start <- current_tour[1]
  n <- length(current_tour)
  all_nodes <- seq(1,dim(tsp)[1],1) #list all nodes
  max <- length(all_nodes)
  max <- min(max, max_length)
  
  while( n < max){
  
    insertion_cost <- c()
    
    # node k is to be inserted
    k <- nearest_insertion(tsp,current_tour)
    
    # get insertion cost between each consecutive nodes i, j
    for (i in current_tour[1:n-1]){             
        j <- current_tour[match(i,current_tour)+1]
        insertion_cost <- c(insertion_cost, distance(tsp,i,k) + distance(tsp,k,j)-distance(tsp,i,j))
    }
    
    #choose best insertion position (which minimizes distance)
    insert_k_pos <- which.min(insertion_cost)+1
    
    #update current tour
    current_tour <- c(current_tour[1:insert_k_pos-1],k,current_tour[insert_k_pos:n])
    n <- length(current_tour)
    
  }
  
  return (c(current_tour,start))
    
}

```

Greedy Algorithm

```{r}

#### Auxiliary functions ####

#return position (i,j) of minimum element in a matrix
min_matrix <- function(M){
  return (arrayInd(which.min(M),dim(M)))
}

#return position (i,j) of maximum element in a matrix
max_matrix <- function(M){
  return (arrayInd(which.max(M),dim(M)))
}

# change vector (i,j) to (j,i)
flip_pair <- function(pair){
  new_pair <- matrix(nrow=1,ncol=2)
  new_pair[1] <- pair[2]
  new_pair[2] <- pair[1]
  return(new_pair)
}

# retrieve the sequence from matrix representation starting from "start"  
nodes_in_sequence <-function(matrix, start){
  s <- c(start)
  #retrieve the sequence
  if (max(matrix[start,],na.rm=T)== -Inf) { return (1)} else {
  while (max(matrix[tail(s,n=1),],na.rm=T) != 0 ){
    s <- c(s,which.max(matrix[tail(s,n=1),]))
    if (length(s) > 1 & tail(s,n=1) == start) { break } # end because we have a cycle
    if(max(matrix[tail(s,n=1),],na.rm=T) == -Inf) { break } #end because the node is not connected to any other node
  }
  return(s)
  }
} 

#is the sequence a cycle  ?
is_cycle <-function(matrix, start){
   s <- nodes_in_sequence(matrix,start)
   return(tail(s,n=1) == start)
} 

#### greedy algorithm ####
greedy_algorithm_tour <- function(tsp) {
  
    n <- dim(tsp)[1]
    X <- matrix(nrow=n, ncol=n) #tour representation in a matrix, it will be initialized with NA
    C<- tsp  # distances matrix
    iteration <- 0
    
    for (i in 1:n){
      C[i,i] <- Inf
    }
     start_node <- min_matrix(C)[1] 
     
    while (sum(is.na(X)) > 0){
      
      # find the minimum element in C
      min <- min_matrix(C) # min = (i,j)
      
      # temporary connect i to j
      X[min] <- 1
      
       #check that no cycles occur until we have connected all nodes
     
      oldw <- getOption("warn")
      options(warn = -1)
      if (is_cycle(X,min[1]) & length(nodes_in_sequence(X,min[1])) != n+1) {
      options(warn = oldw)  
         X[min] <- 0
         C[min] <- Inf
      } else {
        
        # update X and C matrices accordingly
        
        X[min] <- 1            #X[i,j] = 1
        C[min] <- Inf
        X[flip_pair(min)] <- 0 #X[j,i] = 0
        C[flip_pair(min)] <- Inf
        
        for(i in 1:n) {
          if(i != min[1]) {  X[i,min[2]] <- 0 
                             C[i,min[2]] <- Inf  }
          if(i != min[2]) {  X[min[1],i] <- 0
                             C[min[1],i] <- Inf}
        }
      }
      iteration <- iteration + 1
    }
     
    #retrieve tour
    return(nodes_in_sequence(X,start_node))
}

#greedy_algorithm_tour(tsp2)
```


2-Opt algorithm.
```{r}

############################### Auxiliary functions######################################

#  matrix_tour C++ function

#  TwoOptSwap C++ function

# Is a pair of edges consecutive in a given tour?
consecutive_edges <- function(tour, edge1, edge2){
  return ( which(tour==edge1[1])[1] == which(tour==edge2[2])[1] | which(tour==edge1[2])[1] == which(tour == edge2[1])[1])
}

# Node after node1
next_node <- function(tour, node1){
  return(tour[(which(tour==node1)[1]+1)])
}

# Node before node1
previous_node <- function(tour, node1){
  if (length(which(tour==node1))== 1) {
  return(tour[(which(tour==node1)[1]-1)])} else {
    return(tour[(which(tour==node1)[2]-1)])
  }
}

# neighborhood of an edge, if edge is (N(i),N(i+1)), neighborhood is (N(i-1),N(i),N(i+1),N(i+2))
# i index, N(i) element in the position i of the tour

neighborhood <- function(tour, edge){
  n<- length(tour)
  pos1 <- which(tour==as.numeric(edge[1]))
    pos2 <- which(tour==as.numeric(edge[2]))
    if ( length(pos1) == 2) {
      pos1_1  <- pos1[1]
      pos1_2 <- pos1[2]} else {
      pos1_1 <- pos1
      pos1_2 <- 2*n}
    if ( length(pos2) == 2) {
       pos2_1  <- pos2[1]
       pos2_2 <- pos2[2]} else {
      pos2_1 <- pos2
      pos2_2 <- 2*n}
    
    previous_node_pos <- which(tour == previous_node(tour,as.numeric(edge[1])))[1]
    next_node_pos <- which(tour== next_node(tour,as.numeric(edge[2])))[1]
    return(c(pos1_1, pos1_2, pos2_1, pos2_2, previous_node_pos, next_node_pos))
}


############################## 2-opt algorithm ################################

#it starts with a given tour

improvement_2opt <- function(tsp, current_tour, asymmetric, greedy) {
  
  if(missing(asymmetric)) {
    asymmetric <- FALSE
  }
  if(missing(greedy)){
    greedy <- TRUE
  }
  
  df <- tour_to_df(tsp,current_tour)
  df <- df[order(-df$distance),]  # descending order by distance
  df <- subset(df, distance != 0) #reduce the search avoiding edges with 0 distance
  n <- length(current_tour)-1
  starting_min_distance <- tour_length(tsp, current_tour)
  min_distance <- starting_min_distance 
  
  #Vector F and B: cumulative sum of distances when walking the tour forward and backward
  #initialize F and V vector
  
  if(asymmetric){
  f <- cumsumF(tsp,current_tour)
  b <- cumsumB(tsp,current_tour)
  }
  
  it <- 0
  
  improvements <- 0
  unsucessfull_search <- 0
  
  continue <- TRUE #boolean flag
  
  while(continue){
    
    #start with the largest edge in the current tour
    edge1 <- df[1,c("from","to")]
    df <- df[-1,] #to drop first row
    
    it2 <- 0
    neighborhood_edge1 <- neighborhood(current_tour,edge1)
    
    #search for a better tour
    if(asymmetric){
      if(greedy){
         new_tour <- TwoOpt_New_Tour_A(neighborhood_edge1, current_tour, tsp, min_distance, f, b)
      } else{
        new_tour <- TwoOpt_New_Tour_A_v2(neighborhood_edge1, current_tour, tsp, min_distance, f, b)
      }
    } else{
    new_tour <- TwoOpt_New_Tour(neighborhood_edge1, current_tour, tsp, min_distance)
    }
    #improvement?
    if ((length(new_tour) == (n+2)) && (new_tour[n+2] >1e-10 )){  #new_tour[n+2] is the gain, if there's any
      current_tour <- new_tour[1:(n+1)]
      min_distance <- min_distance - new_tour[n+2]
      
      df <- tour_to_df(tsp,current_tour)
      df <- df[order(-df$distance),]  # descending order by distance
      df <- subset(df, distance != 0) #reduce the search avoiding edges with 0 distance
      
      if(asymmetric){
        f <- cumsumF(tsp,current_tour)
        b <- cumsumB(tsp,current_tour) 
      }
      
      improvements <- improvements + 1
      cat("Number of improvements: ", improvements, "\n")
      cat("Tour distance: ",min_distance, "\n" )
      #unsucessfull_search <- 0
    } else {
      unsucessfull_search <- unsucessfull_search + 1
    }
    
    #stopping criterion, when distance is improved in at least 30% abs(min_distance-starting_min_distance)/starting_min_distance > .3 |
      if (  nrow(df)== 0) {
         cat("unsucessfull search: ",unsucessfull_search , "\n")
         continue = FALSE }
  
  }
    return(current_tour)
}

```

3 opt

```{r}

# exhaustive search
ThreeOpt_R<- function (current_tour, dis_matrix, greedy, max_improvements, min_max_gain){
  
  if (missing(greedy)) { greedy <- FALSE}
  if (missing(max_improvements)) { max_improvements <- Inf}
  if(missing(min_max_gain)) { min_max_gain <- 0}
  
  gain <- 1
  improvements <- 0
  
  while((gain > 1e-10) & (improvements < max_improvements)){
    
    if(!greedy) {
      v <- ThreeOpt_BestSwap(current_tour,dis_matrix)
    } else {
      v <- ThreeOpt_GreedySwap(current_tour,dis_matrix, min_max_gain = min_max_gain)  
    }
    current_tour <- v[-length(v)]
    gain <- v[length(v)]
    improvements <- improvements + 1
  }
  return(current_tour)
}

# greedy version of 3 opt
ThreeOpt_R_greedy <- function(current_tour, dis_matrix, max_improvements, nodes_to_avoid){
  
  if(missing(nodes_to_avoid)){nodes_to_avoid <- NULL}
  df <- tour_to_df(dis_matrix,current_tour)
  df <- df[order(-df$distance),]  # descending order by distance
  df<-  subset(df, ! from %in% nodes_to_avoid)
  df<-  subset(df, ! to %in% nodes_to_avoid)
  df <- subset(df, distance != 0) #reduce the search avoiding edges with 0 distance
  
  improvements <- 0
  continue <- TRUE
  
  while(continue){
    
    current_tour <- reorder_tour(current_tour,df$from[1]) #start from the largest edge
    df <- df[-1,]
    
    #search for a better tour
    v <- ThreeOpt_BestSwap_start(current_tour,dis_matrix) #this function returns the best tour obtained by swapping the first edge of the tour and 2 other edges
    gain <- v[length(v)]
    
    if(gain > 1e-10){ #if there is an improvement
      improvements <- improvements + 1
      current_tour <- v[-length(v)] #update current tour
      df <- tour_to_df(dis_matrix,current_tour)
      df <- df[order(-df$distance),]  # descending order by distance
      df<-  subset(df, ! from %in% nodes_to_avoid)
      df<-  subset(df, ! to %in% nodes_to_avoid)
      df <- subset(df, distance != 0) #reduce the search avoiding edges with 0 distance
      }
    
    #stopping criterion
    if(nrow(df)== 0 | improvements == max_improvements){ continue <- FALSE
     cat("improvements: ",improvements , "\n")}
  }
  return(current_tour)
}

```


Plot Tour 

```{r}
library(dplyr)
library(ggplot2)  

plot_tour <- function(coordinates, tour,  labels){
  
  if(missing(labels)){ labels = FALSE}

  m <- coordinates[tour %>% as.numeric,]
  m <- data.frame(x = m$x_coord[1:(nrow(m) -1)], 
                  y = m$y_coord[1:(nrow(m) -1)],
                  xend = m$x_coord[2:nrow(m)],
                  yend = m$y_coord[2:nrow(m)])
 p <- ggplot(data = coordinates, aes(x = x_coord, y =y_coord)) + geom_point() +
             geom_segment(data = m, aes(x = x, xend = xend, y = y, yend = yend)) 
 
 if(labels) {
   p <- p + geom_text(label=rownames(coordinates), nudge_x = 2, nudge_y = 2, check_overlap = T)
 }
  p
}
```




